---
title: "00 Overall Variant-Calling Resource"
author: "subhash"
date: "2026-04-28"
output: html_document
params:
  repo_root: "/home/jyanglab/subhash/sorghum-variant-calling-resource"
  pf_csv: "/Users/subhashmahamkali/Downloads/pf_barcode_reads_summary_v2.csv"
  book1_xlsx: "/Users/subhashmahamkali/Documents/variant_calling/Book1.xlsx"
  hcc_new_root: "/work/jyanglab/subhash/NEW"
  hcc_bqsr_root: "/work/jyanglab/subhash/BQSR"
  hcc_gdb_root: "/work/jyanglab/subhash/6.genomicsdb_5mb"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

repo_root <- normalizePath(params$repo_root, mustWork = FALSE)

read_id_list <- function(path) {
  if (!file.exists(path)) return(character())
  x <- trimws(readLines(path, warn = FALSE))
  unique(x[nzchar(x)])
}

safe_count <- function(x) length(unique(x))
```

# Purpose

This document provides a single, reproducible summary of the full sorghum
variant-calling resource, including:

- sample accounting transitions
- step-wise pipeline organization
- input/output path provenance
- manuscript-ready run order
- current HCC execution snapshot

# Source Files (Cited)

```{r source-files}
source_files <- data.frame(
  category = c(
    "Raw NEW run summary",
    "Curated transition workbook",
    "Final 922 list",
    "Core 876 list",
    "ExPVP 46 list",
    "Transition audit workbook",
    "Included IDs workbook",
    "Removed IDs workbook",
    "HCC manifest path registry",
    "HCC job-chain snapshot"
  ),
  path = c(
    params$pf_csv,
    params$book1_xlsx,
    file.path(repo_root, "metadata/curated/final_922_samples.txt"),
    file.path(repo_root, "metadata/curated/bqsr_876_samples.txt"),
    file.path(repo_root, "metadata/curated/expvp_46_samples.txt"),
    file.path(repo_root, "metadata/curated/sample_transition_audit.xlsx"),
    file.path(repo_root, "metadata/curated/sample_ids_included.xlsx"),
    file.path(repo_root, "metadata/curated/sample_ids_removed.xlsx"),
    file.path(repo_root, "metadata/locations/hcc_manifest_paths.tsv"),
    file.path(repo_root, "metadata/run_tracking/2026-04-28_slurm_submission_chain_steps06_10.tsv")
  ),
  stringsAsFactors = FALSE
)
source_files
```

# Reproducible Sample Accounting

## Repo-curated counts

```{r repo-counts}
final_922 <- read_id_list(file.path(repo_root, "metadata/curated/final_922_samples.txt"))
bqsr_876 <- read_id_list(file.path(repo_root, "metadata/curated/bqsr_876_samples.txt"))
expvp_46 <- read_id_list(file.path(repo_root, "metadata/curated/expvp_46_samples.txt"))

repo_counts <- data.frame(
  metric = c("final_922", "bqsr_876", "expvp_46", "bqsr_plus_expvp"),
  value = c(safe_count(final_922), safe_count(bqsr_876), safe_count(expvp_46), safe_count(union(bqsr_876, expvp_46))),
  stringsAsFactors = FALSE
)
repo_counts
```

# Pipeline Steps and Artifacts

```{r pipeline-table}
pipeline_table <- data.frame(
  step = sprintf("%02d", 1:10),
  name = c(
    "CRAM to FASTQ",
    "fastp trimming",
    "bwa-mem2 alignment",
    "Picard MarkDuplicates",
    "BQSR + HaplotypeCaller",
    "GenomicsDBImport",
    "GenotypeGVCFs",
    "Gather raw VCF",
    "Filter SNPs",
    "Filter indels"
  ),
  script = c(
    "scripts/01_cram_to_fastq/01_prepare_cram_list_and_submit.sh",
    "scripts/02_fastp_trim/fastp_unpaired_array.sh",
    "scripts/03_bwa_mem2_alignment/alignment_unpaired_array.sh",
    "scripts/04_picard_markduplicates/picard_markduplicates_array.sh",
    "scripts/05_bqsr_haplotypecaller/BQSR_array.sh",
    "scripts/06_genomicsdb_import/genomicsdb_import_array_922.sbatch",
    "scripts/07_genotype_gvcfs/genotype_from_gdb_array.sbatch",
    "scripts/08_gather_raw_vcf/gather_raw_vcf.sbatch",
    "scripts/09_filter_snps/filter_snps.sbatch",
    "scripts/10_filter_indels/filter_indels.sbatch"
  ),
  wrapper = c(
    "scripts/manuscript_resource/03_step01_cram_to_fastq_submit.sh",
    "scripts/manuscript_resource/04_step02_fastp_submit.sh",
    "scripts/manuscript_resource/05_step03_alignment_submit.sh",
    "scripts/manuscript_resource/06_step04_picard_submit.sh",
    "scripts/manuscript_resource/07_step05_bqsr_haplotypecaller_submit.sh",
    "scripts/manuscript_resource/08_step06_genomicsdb_import_submit.sh",
    "scripts/manuscript_resource/09_step07_genotype_gvcfs_submit.sh",
    "scripts/manuscript_resource/10_step08_gather_raw_vcf_submit.sh",
    "scripts/manuscript_resource/11_step09_filter_snps_submit.sh",
    "scripts/manuscript_resource/12_step10_filter_indels_submit.sh"
  ),
  stringsAsFactors = FALSE
)
pipeline_table
```

# HCC Submission Snapshot (Steps 06-10)

```{r hcc-submission-snapshot}
snap_file <- file.path(repo_root, "metadata/run_tracking/2026-04-28_slurm_submission_chain_steps06_10.tsv")
if (file.exists(snap_file)) {
  read.delim(snap_file, stringsAsFactors = FALSE)
} else {
  data.frame(note = "Snapshot file not found at expected path")
}
```

# End-to-End Run Order (Manuscript Resource)

```{bash, eval=FALSE}
cd /home/jyanglab/subhash/sorghum-variant-calling-resource/scripts/manuscript_resource
bash 01_prepare_hcc_directories.sh
bash 02_build_manifests.sh
bash 03_step01_cram_to_fastq_submit.sh
bash 04_step02_fastp_submit.sh
bash 05_step03_alignment_submit.sh
bash 06_step04_picard_submit.sh
bash 07_step05_bqsr_haplotypecaller_submit.sh
bash 08_step06_genomicsdb_import_submit.sh
bash 09_step07_genotype_gvcfs_submit.sh
bash 10_step08_gather_raw_vcf_submit.sh
bash 11_step09_filter_snps_submit.sh
bash 12_step10_filter_indels_submit.sh
```
